---
title: "Comparison of USGS and NEON elevation data"
author: "Nate Looker"
date: "June 22, 2016"
output: html_document
---

## Load libraries and define path to GRSM data

```{r load-lib, message=FALSE}

library(raster)
library(rgdal)
library(maptools)

direc <- "/media/look/AOP-NEON1-4/D07/GRSM/2015/"
f_usgs <- "../NEONdata/GRSM_DEM_USGS/grdn36w084_13/w001001.adf"  
```

```{r}

grsm_kml <- getKMLcoordinates(kmlfile=paste(direc, "GRSM_L1/GRSM_Lidar/BoundaryKMLs/full_boundary.kml", sep=""), ignoreAltitude=T)

```


## Read in USGS elevation data for GRSM site

```{r read-usgs-dem}

dem_usgs <- raster(f_usgs)

```

## Read in site spatial extent data (converted from kml using GDAL in bash)

ogr2ogr -f 'ESRI Shapefile' ~/Documents/data/NEONDI-2016/NEONdata/D07_Lidar_Val/grsm.shp /media/look/AOP-NEON1-4/D07/GRSM/2015/GRSM_L1/GRSM_Lidar/BoundaryKMLs/full_boundary.kml


```{r read-boundary-shp}

grsm_boundary <- shapefile("~/Documents/data/NEONDI-2016/NEONdata/D07_Lidar_Val/grsm.shp", stringsAsFactors=F)

plot(dem_usgs)
plot(grsm_boundary, col=2, add=T)
```

## Crop DEM by site boundary and then reproject to CRS of NEON LiDAR data

```{r mask-dem}
# reproject site boundary to same CRS as USGS DEM
grsm_boundary_grs80 <- spTransform(grsm_boundary, CRS(projection(dem_usgs)))

# crop USGS DEM by site boundary
grsm_dem_usgs <- crop(dem_usgs, extent(grsm_boundary_grs80))

newproj <- projection(raster("/media/look/AOP-NEON1-4/D07/GRSM/2015/GRSM_L3/GRSM_Lidar/DTM/2015_GRSM_1_257000_3951000_DTM.tif"))

# reproject clipped USGS DEM to same CRS as NEON LiDAR data
grsm_dem_usgs_utm <- projectRaster(grsm_dem_usgs, crs=newproj)

plot(grsm_dem_usgs_utm)
```

## Export processed USGS DEM for GRSM site

```{r dem-geotiff}

writeRaster(grsm_dem_usgs_utm, filename = "GRSM_DEM_USGS_UTM.tif", 
            format = "GTiff",
            overwrite = T,
            NAflag = -9999)

```


